%% ----------------------------------------------------------------------------
% CVG SA/MA thesis template
%
% Created 03/08/2024 by Tobias Fischer
%% ----------------------------------------------------------------------------

\newpage
\chapter{Related Work}

The related work section has the following purposes: 

\begin{itemize}
 \item \textit{Is the overview concise?} Give an overview of the most relevant work to the needed extent. Make sure the reader can understand your work without referring to other literature.
 \item \textit{Does the compilation of work help to define the ``niche'' you are working in?} Another purpose of this section is to lay the groundwork for showing that you did significant work. The selection and presentation of the related work should enable you to name the implications, differences, and similarities sufficiently in the discussion section.
\end{itemize}

description of task
\section{Visual Question Answering Task (indoor + other)}

VQA vs 3D-QA

indoor is a subtask but it is interesting because we can include some 3D element to it
- also this lays the foundation for answering harder questions later on, which cannot be answered simply with 2D methods: eg. about distances and paths, which would be interesting for navigation tasks.
\begin{itemize}
    \item Task description
    \item Expected inputs / outputs
    \item Question ontology
    \item Classical methods (non-SG), strenghts and limitations
    \item Scene Graph methods (Graph Question Answering), strengths and limitations (most of these for indoor scenes say they use graphs but they never utilise edge info)
\end{itemize}

description of data
\section{Scene Graphs}
\begin{itemize}
    \item Scene graph formal definition + advantages
    \item Scene graph generation methods
    \item Graph neural networks
    \item Applications and limitations
\end{itemize}


description of the "building blocks" used to create the model
\section{Image-text pretrained models}

\section{Transformer encoder / decoders}
GNN can be formulated as a 

\section{Existing datasets}

